\documentclass{article}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{enumitem}
\usepackage{listings}
\newcommand{\bld}[1]{\boldsymbol{#1}}

\begin{document}

\title{MATH520 Exam 3}
\author{Fernando}
\date{\today}
\maketitle

\section*{Problem 1}
\subsection*{Part b}
Since they say the tableau is canonical and we don't know the values of ``?"
the only way we can ensure this is if the basis is $[a_3,a_7,a_2]$, which forces $\beta=0$.
\subsection*{Part a}
From the previous part we know that the basis is $[a_3,a_7,a_2]$, and since the
tableau is canonical we know that the rightmost column corresponds to the
values of the basic variables so in particular $x_7=\alpha$. Since the lower
right corner represents the value of the objective function and we are on phase
I, then it means that all of our artificial variables are 0, in particular
$x_7=0$. So $\alpha=0$.
\subsection*{Part c}
Again, since the tableau is canonical and the lower right corner is 0 it means
that the artificial problem is solved, so the lowermost row must be non
negative. In particular $\gamma\geq0$, so $\gamma=2$.
\subsection*{Part d}
Since the artificial problem is solved, proposition 16.1 tells us that we can
solve the original problem, that means we should be able to bring $a_4$ into
the basis, but if we pivot around $(1,4)$ notice that in the lower right corner
we would end up with $-6\delta$, since this should still be 0 we conclude that
$\delta=0$.
\section*{Problem 2}
\subsection*{Part i}
This is possible. An example is given in the book. Take as the primal
\[
\begin{aligned}
\text{minimize}\quad & x\\
\textrm{s.t.} \quad &x\leq 1,\\
\end{aligned}
\]
then the dual is
\[
\begin{aligned}
\text{maximize}\quad & \lambda\\
\textrm{s.t.} \quad &\lambda= 1\\
\quad &\lambda\leq 0.\\
\end{aligned}
\]
Clearly the primal has a feasible solution and the dual does not.
\subsection*{Part ii}
This is impossible by the duality theorem (if one has an optimal feasible
solution, then so does the other).
\subsection*{Part iii}
This is impossible by the duality theorem (if one has an optimal feasible
solution, then so does the other).
\subsection*{Part iv}
This is possible. An example is given in the book. Take as the primal
\[
\begin{aligned}
	\text{minimize}\quad & [1,-2]x\\
\textrm{s.t.} \quad &\begin{bmatrix}
	1 & -1\\
	-1 & 1
\end{bmatrix}x\geq \begin{bmatrix}
2\\	
-1
\end{bmatrix}.\\
	      \quad &x\geq 0.\\
\end{aligned}
\]
Which has no feasible solution because it requires $[1,-1]x\geq 2$ and
$[1,-1]x\leq 1$. The dual of this problem is
\[
\begin{aligned}
	\text{maximize}\quad & \lambda^T\begin{bmatrix}
	2\\	
	-1
	\end{bmatrix}\\
\textrm{s.t.} \quad &\lambda^T\begin{bmatrix}
	1 & -1\\
	-1 & 1
	\end{bmatrix}\leq [1,-2]\\
	      \quad &\lambda\geq 0.\\
\end{aligned}
\]
Which again has no feasible solution because it requires $\lambda^T\begin{bmatrix}
1\\
-1
\end{bmatrix}\leq 1$ and $\lambda^T\begin{bmatrix}
1\\
-1
\end{bmatrix}\geq 2$.
\section*{Problem 3}
\subsection*{Part a}
Our problem is the following
\[
\begin{aligned}
\text{minimize}\quad & f(x)\\
\textrm{s.t.} \quad &g(x)\leq 0.\\
\end{aligned}
\]
Where $f(x)=x_1x_2$ and $g(x)=\begin{bmatrix}
g_1(x)\\
g_2(x)
\end{bmatrix}=\begin{bmatrix}
2-x_1-x_2\\
x_1-x_2
\end{bmatrix}$

Since in this problem we only have inequality constraints, the KKT condition
is:
\begin{enumerate}[I]
\item $\mu^*\geq0$ \label{muPositive}
\item $Df(x^*)+{\mu^*}^TDg(x^*)=0^T$ \label{KKT}
\item ${\mu^*}^Tg(x^*)=0$ \label{muTg}
\item $g(x^*)\leq0$ \label{gNegative}
\end{enumerate}
\subsection*{Part b}
Notice that condition \ref{muTg} can be written as
\begin{equation} \label{cond3reWrite}
\mu_1g_1 + \mu_2g_2=0.
\end{equation}
Because of conditions \ref{muPositive} and \ref{gNegative} each of the
terms in (\ref{cond3reWrite}) must be 0, so $\mu_ig_i=0$.
If we suppose $\mu_1=0$ condition \ref{KKT} gives us the following:
\begin{align*}
0^T&=Df(x^*)+{\mu^*}^TDg(x^*)\\
&=[x_2,x_1]+[\mu_1,\mu_2]\begin{bmatrix}
-1 & -1\\
1 & -1
\end{bmatrix}\\
&=[x_2,x_1]+[0,\mu_2]\begin{bmatrix}
-1 & -1\\
1 & -1
\end{bmatrix}\\
&=[ x_2+\mu_2, x_1-\mu_2].
\end{align*}
Adding these two equations we get
\[
x_1+x_2=0,
\]
but this is a contradiction with condition \ref{gNegative}, in particular it
contradicts $g_1(x)=2-x_1-x_2\leq0$. Conclusion: $u_1\neq 0$, so we know that
for sure $g_1=0$.
With regards to the other term in (\ref{cond3reWrite}), if we assume
$g_2(x)=0$, this means $x_1=x_2$ and replacing this in \ref{KKT} we
get
\begin{align*}
0^T&=Df(x^*)+{\mu^*}^TDg(x^*)\\
&=[x_1,x_1]+[\mu_1,\mu_2]\begin{bmatrix}
-1 & -1\\
1 & -1
\end{bmatrix}\\
&=[ x_1-\mu_1+\mu_2, x_1-\mu_1-\mu_2].
\end{align*}
Again adding these equations we get $\mu_1=x_1$, and then $\mu_2=0$.
Conclusion: we know that for sure $\mu_2=0$. Once again using
\ref{KKT} we get:
\begin{align*}
0^T&=Df(x^*)+{\mu^*}^TDg(x^*)\\
&=[x_1,x_2]+[\mu_1,\mu_2]\begin{bmatrix}
-1 & -1\\
1 & -1
\end{bmatrix}\\
&=[x_1,x_2]+[\mu_1,0]\begin{bmatrix}
-1 & -1\\
1 & -1
\end{bmatrix}\\
&=[ x_1-\mu_1, x_2-\mu_1].
\end{align*}
From where $x_1=x_2=\mu_1$ and using the fact that $g_1=0$ we get that
$x_1=x_2=\mu_1=1$.

So the only point that satisfies the KKT condition is ${x^*}^T=[1,1]$, with the
KKT multiplier vector ${\mu^*}^T=[1,0]$.

Since $\nabla g_1(x^*)=\begin{bmatrix} -1\\-1\end{bmatrix}$ and $\nabla
g_2(x^*)=\begin{bmatrix} 1\\-1\end{bmatrix}$ are L.I. $x^*$ is regular.
\subsection*{Part c}
First we need to determine $T(x^*)$. Notice that both constrains are active so
\[
	T(x^*)=\{y\in \mathbb{R}^2:Dg(x^*)y=0\}
\]
and since $Dg(x^*)$ is invertible we have
\[
	T(x^*)=\{0\},
\]
so the SONC is trivially satisfied.
\subsection*{Part d}
In this case we have
\[
	\tilde{J}(x^*,\mu^*)=\{1\},
\]
and thus
\[
	\tilde{T}(x^*,\mu^*)=\{y:[-1,-1]y=0\}=\text{span}([1,-1]^T).
\]
Now we compute $F(x)$ and see that
\[
	F(x)=\begin{bmatrix}
		0 & 1\\	
		1 & 0
	\end{bmatrix},
\]
also notice that because $g_1$ and $g_2$ are linear in each variable we have
\[
G_1(x)=G_2(x)=\begin{bmatrix}
	0 & 0\\	
	0 & 0
\end{bmatrix},
\]
then
\[
L(x,\mu)= F(x)=\begin{bmatrix}
		0 & 1\\	
		1 & 0
	\end{bmatrix},
\]
and notice that $[1,-1]^T\in\text{span}([1,-1]^T)=\tilde{T}(x^*,\mu^*)$, and
\[
	[1,-1]L(x^*,\mu^*)\begin{bmatrix}
1\\	
-1
\end{bmatrix}=
[1,-1]
\begin{bmatrix}
0 & 1\\	
1 & 0
\end{bmatrix}
\begin{bmatrix}
1\\	
-1
\end{bmatrix}=
[-1,1]\begin{bmatrix}
1\\	
-1
\end{bmatrix}=-2<0.
\]
So no point in part c satisfies the SOSC.
\end{document}
